---
title: "Clinical Trials Part 2"
author: "David Andai"
date: "11/30/2022"
output: github_document
---
Clinical trials can be expensive to run. Due to the high cost of running one, Big Pharma conduct most of the clinical trials. In a couple of instances, Big Pharma has been in the spotlight for influencing the results of clinical trials. A negative influence on clinical trial results and outcomes may be dangerous to the population.

The scope of Clinical trial sponsors has actively changed over a couple of years to include independent organizations and professionals, following designs and protocols agreed upon by all parties and approved by ethical review boards. 
Long-term phase 3 trials are vital to sponsors and aimed at providing an adequate basis for a submission for marketing approval. These studies have a long duration and are likely to show long-term or rare side effects. Studies in phase 3 are conducted across a large patient population and should include a population representative of the intended target population that will use the drug. Hence, phase 3 trials can be multicenter, occurring in different countries. 
Here are some of the Sponsors of Phase 3 clinical trials that are ongoing in Kenya.
World Health Organization
Title: Low dose magnesium sulphate for treatment of seizures in pregnancy
Target size:
Condition/Disease: Pre-Eclampsia
Drug: Magnesium Sulphate
Estimated completion date: 2023
Global Network for Womens' and Childrens' Health
Title: Prevention of Iron Deficiency Anemia Post-delivery (PRIORITY)
Target size:
Condition/Disease:  Postpartum anemia
Drug: Iv iron infusion, Oral tablets
Estimated completion date: December 1, 2025
Hoffman La Roche
Title: Phase-III STUDY TO DETERMINE SAFETY AND EFFICACY  OF GIREDESTRANT COMPARED WITH PHYSICIANS CHOICE IN PATIENTS WITH ESTROGEN RECEPTOR POSITIVE, HER2 NEGATIVE EARLY BREAST CANCER
Study Design: Randomized, Parallel Group, Active Controlled Trial
Condition/Disease: Breast Cancer
Drug: Giredestrant, Endocrine Therapy of Physician's Choice, LHRH Agonist
Estimated completion date: December 19, 2025

Academic Medical Centre at the University of Amsterdam
Title: Evaluation of a new and simple diagnostic test for malaria
Target size: 4700
Condition/Disease: Malaria
Drug: Rapid Diagnostic Test
Estimated completion date: 2023
Global Antibiotic Research and Development Partnership
Title: NeoSep1: a study to determine the ranking of existing and new antibiotics combinations to treat newborn babies who are in hospital with severe sepsis
Target size: 3060
Condition/Disease: Neonatal sepsis
Drug: Amikacin, Flomoxef and Fosfomycin
Estimated completion date: 2023

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("XML")
library(ggplot2)
library(dbplyr)
library(tidyverse)
library(lubridate)
library(ggthemes)
```


```{r include=FALSE, warning=FALSE}
#Load Clinical Trials Data
setwd("C:/Users/user/Documents/R/Clinicals2")

Kenya <- xmlToDataFrame("ICTRP-Results Kenya.xml")
summary(cars)
```


```{r include=FALSE, warning=FALSE}
# get the Phase column
Phase <- Kenya$Phase
#set to lower case
Phase <- tolower(Phase)
#Remove hyphens
Phase <- gsub("-/@#$:", "", Phase )
#Remove \n
Phase <- gsub("\n", "", Phase)


##Phase 4
Phase <- gsub('phase-4','phase 4', Phase)
Phase <- gsub('phase iv','phase 4', Phase)
Phase <- gsub('phase 1v','phase 4', Phase)
Phase <- gsub('iv','phase 4', Phase)

##Phase3
Phase <- gsub('phase-3','phase 3', Phase)
Phase <- gsub('phase iii','phase 3', Phase)


#Phase2
Phase <- gsub('phase-2','phase 2', Phase)
Phase <- gsub('phase ii','phase 2', Phase)
Phase <- gsub('ii','phase 2', Phase)
Phase <- gsub('phase 2i','phase 2', Phase)
Phase <- gsub('phase2/phase 2i','phase 2', Phase)
Phase <- gsub ('phase 2/phase 2','phase 2', Phase)

#Phase1
Phase <- gsub('phase-1','phase 1', Phase)
Phase <- gsub('phase i','phase 1', Phase)


#Phase0
Phase <- gsub('phase-0','phase 0', Phase)
Phase <- gsub('0','phase 0', Phase)
Phase <- gsub('phase phase 0','phase 0', Phase)

#Phase2/3
Phase <- gsub('phase 2/ phase 3','phase 2 / phase 3', Phase)
Phase <- gsub('phase 2/phase 3','phase 2 / phase 3', Phase)

#Phase1/2
Phase <- gsub('phase 1/phase 2','phase 1 / phase 2', Phase)
Phase <- gsub('phase 1/ phase 2','phase 1 / phase 2', Phase)

#Replace the long texts to empty strings
Phase <- gsub('explanatory', '', Phase)
Phase <- gsub('trials','', Phase)
Phase <- gsub('[()]', '', Phase)
Phase <- gsub('human', '', Phase)
Phase <- gsub('pharmacology', '', Phase)
Phase <- gsub('confirmatory', '', Phase)
Phase <- gsub('exploratory', '', Phase)
Phase <- gsub('therapeutic', '', Phase)
Phase <- gsub('use', '', Phase)
Phase <- gsub('no', '', Phase)
Phase <- gsub('n/a', '', Phase)

#Substitute the remaining patterns
Phase <- gsub('phase 1: yes', 'yes1', Phase)
Phase <- gsub('phase 2: yes', 'yes2', Phase)
Phase <- gsub('phase 3: yes', 'yes3', Phase)
Phase <- gsub('phase 4: yes', 'yes4', Phase)

Phase <- gsub('phase 1:', '', Phase)
Phase <- gsub('phase 2:', '', Phase)
Phase <- gsub('phase 3:', '', Phase)
Phase <- gsub('phase 4:', '', Phase)

Phase <- gsub("-","", Phase)
Phase <- gsub("[[:space:]]", "", Phase)

Phase <- gsub("yes1", "phase1", Phase)
Phase <- gsub("yes2", "phase2", Phase)
Phase <- gsub("yes3", "phase3", Phase)
Phase <- gsub("yes4", "phase4", Phase)

Phase <- gsub("/","", Phase)
Phase <- gsub("[[:space:]]", "", Phase)

Phase <- gsub('postmarketingsurveillance', 'phase4', Phase)
Phase <- gsub('tspecified','', Phase)
Phase <- gsub('tapplicable','Not_App', Phase)#not_applicable

Phase <- gsub('3','phase3', Phase)
Phase <- gsub('2','phase2', Phase)

Phase <- gsub('phasephase','phase', Phase)
Phase <- gsub('phase', 'p', Phase)

Kenya["fix_Phase"] <- Phase

```


```{r include=FALSE, warning=FALSE}

#Transform Date Column from character to Date variable
class(Kenya$Date_registration) ##'character'
#using the lubridate library
Kenya <- Kenya %>%
  mutate(Date_registration = lubridate::dmy(Date_registration))
class(Kenya$Date_registration)##'Date'

this_year <- Kenya[Kenya$Date_registration >= "2022-01-01" & Kenya$Date_registration <= "2022-10-31", ]
this_year <- this_year[-c(96,97,98),]

#create count column
count_a <- rep(1,times =95)
this_year <- cbind(this_year, count_a)

#Create A Table for sponsors
Sponsors <- this_year[ , c("fix_Phase","Primary_sponsor", 
                           "Countries","Scientific_title","Target_size", "Public_title")]

###

#Change all empty cells to NA
Sponsors [Sponsors == ""] <- NA
Sponsors <- drop_na(Sponsors)

#Obtain Only Phase 3 Clinical trials
p3 <- Sponsors[Sponsors$fix_Phase =='p3',]
##Change Target size column into numeric data type
p3$Target_size = as.numeric(as.character(p3$Target_size))
class(p3$Target_size)

#Add ID column
#p3 <- tibble::rowid_to_column(p3, "ID")

#Drop duplicate Clinical trials that have been duplicated.
## Drop "Serum Institute of India" Duplicate 
rownames(p3) <- 1:nrow(p3)
p3 <- p3[-c(5),]
rownames(p3) <- 1:nrow(p3)

#ID 7 Primary_sponsor has a long value
p3[7, ]
p3[7, 2]= 'Hoffmann-La Roche'
#Change ID 12 Primary sponsor value
p3[12, 2]= 'Hoffmann-La Roche 2'

## Drop "Academic Medical Center" Duplicate 
p3 <- p3[-c(9),]
rownames(p3) <- 1:nrow(p3)

# Rename MSD
p3[9, 2]= 'Merck Sharp & Dohme LLC'
```

```{r warning=FALSE, message=FALSE}
#Data Visualization: Primary Sponsors of on going Phase 3 Clinical Trials in Kenya
plot1 <- ggplot(data = p3,
       aes((y = reorder(Primary_sponsor, -Target_size)), x = Target_size, fill = Target_size))+
  geom_bar(stat = "identity", show.legend = FALSE)+
  labs (title = "Phase 3 Clinical Trials in Kenya",
        subtitle = "In 01-01-2022 to 31-10-2022",
        y = "Primary Sponsors",
        x = "Population Target size")+
  theme_few()
```

```{r message=FALSE, warning=FALSE, dev='png'}
plot1
```

